---
title: "Chaffinches *(again)*"
author: "Hannah Clark"
date: "04/11/2020"
output: html_document
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      fig.retina = 3)
```

```{r packages}
library(tidyverse)
library (bookdown)
```

## Introduction
Chaffinches, *Fringilla coelebs*, are common to UK woodland, farmland, parks, and gardens [@Woodland_Trust_undated-gv]. The average weight of a chaffinch is 18-29g [@noauthor_undated-dy]. The masses of a group of finches were taken and categorised by sex, to see how this varied from this known average.

## Methods
The chaffinches data was imported and tidied

```{r data tidying, echo = TRUE}
#read in the data
chaff <- read.table("~/R Workshop/chaff02/data-raw/chaff.txt", header = T)

#tidying the data
chaff2 <- gather(data = chaff, key = sex, value = mass)

#write the tidied data into a new file
write.table(chaff2, file = '~/R Workshop/chaff02/data/tidychaff.txt', row.names = FALSE)
```

Summarising the data

<!-- this way doesn't take into account the groups, so is less useful  -->

```{r ss(x) version 1, include = FALSE}
sum_sq <- function(df, measure) {
  df %>% summarise(ss = sum(({{measure}} - mean({{measure}}))^2))
}

sum_sq(chaff2, mass)
```

<!-- this does do it by group -->

```{r ss(x) by group, echo = TRUE}
sum_sq <- function(df, measure, group) {
  df %>% group_by({{ group }}) %>% 
    summarise(ss = sum(({{measure}} - mean({{measure}}))^2))
}

sum_sq(chaff2, mass, sex)

```

```{r mean,length, sd, echo = TRUE}
chaff2 %>% group_by(sex) %>% 
  summarise(mean = mean(mass),
            n = length(n),
            sd = sd(mass),
            se = sd/sqrt(n))
```

This can also be done in 1 function

```{r summary in 1 function, echo = TRUE}
data_summary <- function(df, measure, group) {
  df %>% group_by({{ group }}) %>% 
    summarise(mean = mean({{measure}}),
            n = length(n),
            sd = sd({{measure}}),
            se = sd({{measure}})/sqrt(length(n)),
            ss = sum(({{measure}} - mean({{measure}}))^2))
}

data_summary(chaff2, mass, sex)
```


## Results

The results from summarising the chaffinches data, grouping by sex

```{r results}
data_summary(chaff2, mass, sex)
```


## Discussion











